Localization properties, observable character and computational reliability of the linear response in open systems

### 2018-07-08 AnalyticStructure_main.tex
continuing

\section{Introduction}
Since the days of their foundations, disciplines like Computational Physics or Quantum Chemistry always had to deal with the problem of the 
computational reliability of the results. More specifically, a computational result may be considered as reliable for various reasons.
It might be, for example, in agreement with some previously existing data that the employed computational model is susceptible to reproduce.
The model itself may in this way be classified as ``accurate''; it can then be employed to extract, \textit{in silico},
other quantities that might be then used to \emph{predict} or \emph{interpret} experimental data.
Another concept of reliability is the concept of  ``precision'', which is more focused to the actual results
that a given computational model provides for a specific set of quantities, regardless of their potential agreement with 
experimental data. The aim of precise computational approach is to reduce the \emph{computational} uncertainties
of quantities extracted out of a well-defined model, and provide reference results, to which other computer codes
employing the same model can be compared.

A recent, remarkable example of precision-driven initiative is provided by the DeltaCodes project~\cite{deltaTest2016},
where a grassroot community of code-owners struggled to extract well-defined computational results in the context of 
Density Functional Theory (DFT) calculations. Ground-State quantum mechanical quantities as Lattice constants and bulk modulus of 
elementary crystals extracted from different codes were compared with each other, with the objective of reducing the uncertainties 
on such data, and in some sense, to define a \emph{calibration} procedure for a Solid-State Physics computer code. 
Any other code able to extract the same quantities with the same approximations might be compared with the 
results of the community in order to assess its computational reliability.

@@

Our considerations will also reveal useful for the understanding of the \emph{analytic structure} of LR quantities like
the linear susceptibility of the system (i.e. the reducible polarizability).
Although we refer, as anticipated, to TDDFT models, such considerations might also be transposed to Many-Body Perturbation Theory
approaches, in the context Quasi-Particle equations.

@@

%\section{Localization properties of the linear response actors in open systems}

@@

\subsection{Long range behavior of the eigenvalue solution of the Schr\"odinger equation}
\label{SEopenSystem}
Consider a (one-body) wavefunction $\ket{\psi}$ solution of the equation
$\op H\ket{\psi}=\eps\ket{\psi}$, where we split the Hamiltonian operator into the sum of
a kinetic $\op T$ and a potential $\op V \equiv \op H - \op T$ term.
In view of analysing the impact of the BC, it is instructive to transform 
the Schr\"odinger equation in a scattering problem, by writing
\be\lb{SEHelmholtz1}
\ket{\psi} = \op G_H(\eps)\ket{\op V\psi}\;,
\ee
where we have introduced the Green function of the Helmholtz operator $\op G_H(\eps) = (\eps-\op T)^{-1}$,
resolvent of the free-particle Hamiltonian $H_s \equiv \op T$.

When describing the Ground State of a molecule, we suppose that the potential operator, even if non-local,
can be effectively restricted on wavefunctions that can be considered as nonzero only on a
bounded domain. In other terms, a molecule can be associated to a ``scatterer'' that has a \emph{finite} 
spatial extension; if a state $\ket{\psi}$ is localized, it will be the same for the ket $\ket{\op V \psi}$, though 
perhaps in a larger region.
For a computational analysis of the problem, this property is fundamental in the choice of the numerical treatment as we know that basis functions 
that are tailored to express asymptotically vanishing wavefunctions are better suited for the solution of the problem.

@@

It can be seen that for each choice of $p$, the value $\omega = |\eps_p|$ represents the threshold level 
that govern the localization properties of the corresponding fluctuation state.
As a \emph{stringent} hypothesis, let us now assume that the ket $\ket{s_p^\Phi(\omega)}$ is localized, namely its evaluation can be
restricted (if suitable) on a bounded real-space domain.
%The localization character of fluctuation states can be assessed by following the same argument discussed in section \ref{SEopenSystem} for the eigenstates of the SE. In this case 
%the source term depends also from the perturbation added to the system so this analysis applies only to the class 
%of perturbing field for which $\ket{\op\Phi\psi_p}$ is localized, when projected
%in the $\r$-representation. 
Even when such locality argument holds true, for $\omega > |\eps_p|$, 
the Helmholtz kernel behaves as a spherical wave and  gives raise to a $\ket{f^\Phi_p(\omega)}$ which is \emph{delocalized} in the entire real-space domain.
Such delocalization is an intrinsic consequence of the value of $\omega$, and does not depend on the particular choice of the operators.
Only for $\omega < |\eps_p|$ the kernel contains an exponential damping factor and $\ket{f^\Phi_p(\omega)}$ exhibits a bound-state behavior in the long range, 
once again provided that the above locality arguments on $\ket{s_p^\Phi(\omega)}$ -- hence on $\op V'$ and $\Phi$ -- are valid.

Such locality property of the fluctuation state can be directly confirmed for static perturbations at $\omega=0$.
Let us consider a static perturbing operator $\op\Phi_s$ that can be restricted to a finite domain. 
Any local operator, like \emph{e.g.} a static electric field $\mathbf F$, with associated scalar potential $\Phi_s(\r) = -\mathbf F \cdot \r$, would satisfy such property.
We consider the perturbed hamiltonian $\hnot[\dm_{\Phi_s}] + \op\Phi_s$, where $\dm_{\Phi_s}$ represents the GS density of the perturbed system, and assume that the 
perturbation strength ($|\mathbf F|$ in the above example) is sufficiently small to be consistent with a LR treatment of the problem. 
In this case, the occupied eigenfunction $\ket{\psi_p^{\Phi_s}}$ of the perturbed hamiltonian can be extracted with traditional GS techniques, as by the hypothesis 
of Eq.~\eqref{SEHelmholtz1}, they exhibit bound-state behavior.\footnote{The fulfilling of Eq.~\eqref{SEHelmholtz1} for energies below threshold
is not \textit{per se} a guarantee of a localized $\ket{\psi}$. With such equation we demonstrate that, \emph{in the case when} a localized
solution exists, this cannot be found for energies above threshold.}
\begin{figure}[t]
\includegraphics[scale=0.68]{CO_statPolvsBox.eps}
\caption{\label{co_alphaStatic}(Color online) Convergence of the statical polarizability (blue continuous line) and of the ground state permanent dipole (red dotted line) 
of the $CO$ molecule with respect to the size of the computational domain. The inset plot evidences the relative percentage convergence using the value associated to the largest 
box as reference.}
\end{figure}

@@

This fact has interesting implication in the evaluation of LR quantities. Indeed, writing the response 
density in terms of fluctuation states implies that 
\be\lb{LinearResponseFunctDef2}
\braket{\delta\op O}_\Phi(\omega) = 2\mathrm{Re} \left(\sum_p \bra{\psi_p}\op O\ket{f_p^\Phi(\omega)}\right) \;.
\ee
%so the computation of the linear response functional passes through the evaluation of the scalar products indicated in \eqref{LinearResponseFunctDef2}.
Let us suppose that also the observable of interest $\op O$ exhibit the locality properties of the potentials $ \op V$, $\op \Phi$, $\op V'$.
Once again, it would be enough to consider a local operator for this condition to be met.
We observe that there exists two 
distinct regimes, on the basis of value of $\omega$. 
The first one, which we will refer to as the \emph{below threshold} regime, is realized when
$\omega$ is lower than the ionization potential $|\eps_h|$.
In this case all the fluctuation states are below their threshold level and equation \eqref{LinearResponseFunctDef2} 
is expressed in terms of genuinely localized 
quantities. The quantities $\braket{\psi_p \op O| f_p^\Phi(\omega)}$, 
being scalar products of bound-state-like objects, may thus be evaluated with computational setup that
are similar to the ones usually employed for GS calculations.
This consideration proves why, when a given computer code employs established GS numerical techniques, 
\emph{it is much easier to converge LR quantities for $\omega$ below threshold}.
The convergence in this regime is ``strong'', in the sense that the basis functions employed enable to express 
with arbitrary precision \emph{both} the states $\ket{\op O \psi_p}$ and $\ket{f_p^\Phi(\omega)}$.
This is of course true even when the above states are only implicitly expressed by the employed LR treatment,
 as e.g. in the case of Krylov spaces generated from the states $\ket{\op\Phi(\omega) \psi_p}$ \cite{baroni2006,baroni2008,linlinKPM}.

@@

\subsection{The analytic structure of linear susceptibility}
Being themselves transverse operators, without loss of generality we can parametrize excitations by employing  
the \emph{excited states} $\ket{\phi^a_p}$ and $\bra{\chi^a_p}$ that, by construction, are orthogonal to the occupied orbitals
$\ket{\psi_p}$ and satisfy the normalization condition % and the orthonormalization condition \eqref{orthoExcitatioOpDef1} implies 
\be\lb{ExcitedStateOrthNormDef1}
\sum_p \left(\brket{\phi_p^a}{\phi_p^b} - \brket{\chi_p^b}{\chi_p^a}\right) = \delta_{ab} \;. 
\ee
In this way we can parametrize the excitation operators of the system as:
\begin{align}\lb{ExcitationOperatorsDef2}
\op E_a &= \sum_p \left( \ket{\phi^a_p}\bra{\psi_p} + \ket{\psi_p} \bra{\chi^a_p}\right)\;, \nn \\
\op{\tilde E}_a = \commutator{\dmnot}{\op E^t_a} &= \sum_p \left( \ket{\psi_p}\bra{\phi^a_p} - \ket{\chi^a_p} \bra{\psi_p}\right)\;.
\end{align}
Each excitation ``mode'' of the system $\op E_a$ with energy $\Omega_a$, can then be associated to a set of states $\{\phi^a_p,\chi^a_p\}_{p \in a}$,
entirely defined in the unoccupied subspace. For each choice of $p,a$, these objects represent, respectively, the state in which the occupied state $\ket{\psi_p}$
is excited -- or from whom it decays -- when the system is subject to the perturbation $\op E_a^t$.
We denote with the shorthand $p \in a$ the fact that the sum of Eq.~\eqref{ExcitationOperatorsDef2}
is in general restricted only to a $a$-dependent subset of the occupied orbitals, depending on the particular symmetry 
of the excitation.

% We observe that right eigenvalues of $\Liouv$ can be written in terms of the left ones due to the properties 
% of the Liouvillian described in appendix \ref{LiouvillianAction}. 
% By applying equations \eqref{LiouZeroDef1},\eqref{CouplDef1} and \eqref{LiouvillianRightActionDef1} presented in 
% appendix \ref{LiouvillianAction} to the dyadic representations of the excitation 
% operators, 
By applying the equations \eqref{ExcitationOperatorsDef1} to the above parametrization
of the excitations, we obtain the following Sternheimer-like equations for the excited states 
\begin{align}\lb{ExcitationOperatorsDef3}
&\left[\Omega_a-(\hnot - \eps_p)\right] \ket{\phi_p^a} = \op Q_0 \op V'[\op E_a] \ket{\psi_p} \;, \nn\\
&\left[-\Omega_a-(\hnot - \eps_p)\right]\ket{\chi_p^a} = \op Q_0\op V'[\op E^t_a]\ket{\psi_p}  \;.
\end{align}
By writing these equations in a basis set of virtual states we obtain the well-known Casida's eigenvalue equation
associated to the excitation energy $\Omega_a$ (see appendix \ref{casida}).
The excitation spectrum is symmetric w.r.t the inversion of the eigenvalues 
$\Omega_a \rightarrow -\Omega_a$ and the associated excitation with negative energy are described by the transpose pair $\{\chi^a_p,\phi^a_p\}$.
We concentrate our analysis on the sector of positive energies, i.e. $\Omega_a > 0$.

With the same spirit of the previous sections, we present a formal solution of \eqref{ExcitationOperatorsDef3} as follows 
\begin{align}
\ket{\phi^a_p} &= \op G_H(\Omega_a+\eps_p)\ket{s_{\phi}} \;, \nn \\
\ket{\chi^a_p} &= \op G_H(-\Omega_a+\eps_p)\ket{s_{\chi}} \;,
\end{align}
where the source terms $s_\phi$ and $s_{\chi}$ are defined as:
\begin{align}
 \ket{s_{\phi}} &= \op V \ket{\phi^a_p} + \op Q_0 \op V'[\op E_a] \ket{\psi_p} \\
 \ket{s_{\chi}} &= \op V \ket{\chi^a_p} + \op Q_0 \op V'[\op E^t_a] \ket{\psi_p}\;.
\end{align}
We first point out that the Helmholtz kernel associated to the $\ket{\chi^a_p}$ state 
contains an exponential damping factor for \emph{all} the positive values of $\Omega_a$.
The asymptotic long-range behaviour of $\ket{\chi^a_p}$ thus depends only on the locality properties of 
$\ket{s_{\chi}}$, which in turn is related to the behaviour of the operator $\op V'[\op E^t_a]$.
% As in the previous section, we assume the hypothesis that such operator is such that
% the state $\op Q_0 \op V'[\op E^t_a] \ket{\psi_p}$ is localized, at least when \emph{both}
% the states $\phi_p^a$ $\chi_p^a$ have bound-state-like behaviour.

On the other hand, the long-range behaviour of $\op G_H(\Omega_a+\eps_p)$ depends from 
$\Omega_a$, with a threshold of value $\Omega_a=|\eps_p|$. 
For energies above this value, the behaviour of the state $\ket{\phi^a_p}$ is 
delocalized regardless of the particular nature of the operators $\op V$ and $\op V'$.
This means that, if an excitation operator $\op E_a$ has an energy such that
$\Omega_a>|\eps_p|$  for \emph{at least one} of the $p \in a$ defining the excitation,
such operator cannot be restricted to localized states.

@@

@@

We have seen in Sec.~\ref{FluctuationState} that 
for values of $\omega$ below the first ionization potential we expect
$\ket{f_p^\Phi(\omega)}$ to have bound-state like behavior.
For this to be true, the contribution from the delocalized excited states defining
$\ket{f_p^\Phi(\omega)}_C$ should resum into a localized, asymptotically vanishing state
\footnote{We might have locality of $\ket{f_p^\Phi}$ also when perturbation operator $\op \Phi$ is such that 
$f_p^a(\op \Phi,\omega)=0$ for any $a$ labelling the continuum excitations. 
However, this is not true in general: the plot in Fig.~\ref{co_AlphaExc}
show this evidence when $\op \Phi$ is the dipole operator.}.

This means that, contrary to GS, \emph{for LR calculation,
even when the response density is localized, the contribution coming from the 
continuum sector of the excitations can never be neglected}. 
We are not referring here to the continuum eigenstates of the 
unperturbed GS hamiltonian, %which is not in itself important for expressing the linear susceptivity, 
but to the continuum of excitations expressed as eigenstate of the 
Liouvillian superoperator of Eq.~\eqref{Liouvillian}, needed to 
express the spectral representation of the linear susceptibility.
A fluctuation state, even if genuinely localized, will \emph{always} have a nonzero projection
in the essential part of the Liouvillian spectrum.

>>>>

\section{Introduction}
Since the days of their foundations, disciplines like Computational Physics or Quantum Chemistry always had to deal with the problem of the 
computational reliability of the results. More specifically, a computational result may be considered as reliable for various reasons.
It might be, for example, in agreement with some previous observation or measurement that the employed computational model is susceptible to reproduce.
The model itself may in this way be classified as ``accurate''; it can then be employed to extract, \textit{in silico},
other quantities that might be then used to \emph{predict} or \emph{interpret} experiments.
Another concept of reliability is the concept of  ``precision'', which is more focused to the actual results
that a given computational model provides for a specific set of quantities, regardless of their potential agreement with 
experimental data. The aim of precise computational approach is to reduce the \emph{computational} uncertainties
of quantities extracted out of a well-defined model, and provide reference results, to which other computer codes
employing the same model can be compared.

A recent, remarkable example of precision-driven initiative is provided by the DeltaCodes project~\cite{deltaTest2016},
where a grassroot community of code-owners struggled to extract well-defined computational results in the context of 
Density Functional Theory (DFT) calculations. Ground-State quantum mechanical quantities as lattice constants and bulk modulus of 
elementary crystals extracted from different codes were compared with each other, with the objective of reducing the uncertainties 
on such data, and in some sense, to define a \emph{calibration} procedure for a Solid-State Physics DFT code. 
Any other code able to extract the same quantities with the same approximations might be compared with the 
results of the community in order to assess its computational reliability.

@@

The paper is organized as follows.
Initially we will inspect how the question of reproducibility may be expressed 
for ground state calculations of molecules. 
Following these guidelines, we will then move to LR equations,
by first considering the equations of motion of the response density operator, given
the application of a perturbing field.
Then we will consider the behaviour of the ``free oscillations'', i.e. the excited states
of the molecule.
Our considerations will also reveal useful for the understanding of the \emph{analytic structure} of LR quantities like
the linear susceptibility of the system (i.e. the reducible polarizability).
We support our considerations with numerical results on simple LR-TDDFT quantities.
For readability purposes, the details of each calculation are explained in the appendix.

Although we refer, as anticipated, to TDDFT models, such considerations might also be transposed to Many-Body Perturbation Theory
approaches, in the context Quasi-Particle equations.

@@

%\section{Localization properties of the linear response actors in open systems}
\subsection{Long range behavior of the eigenvalue solution of the Schr\"odinger equation}
\label{SEopenSystem}

@@

Consider a (one-body) wavefunction $\ket{\psi}$ solution of the equation
$\op H\ket{\psi}=\eps\ket{\psi}$, where we split the Hamiltonian operator into the sum of
a kinetic $\op T$ and a potential $\op V \equiv \op H - \op T$ term.
In view of analysing the impact of the BC, it is instructive to transform 
the Schr\"odinger equation in a scattering problem, by writing
\be\lb{SEHelmholtz1}
\ket{\psi} = \op G_H(\eps)\ket{\op V\psi}\;,
\ee
where we have introduced the Green function of the Helmholtz operator $\op G_H(\eps) = (\eps-\op T)^{-1}$,
resolvent of the free-particle Hamiltonian $H_s \equiv \op T$.

When describing the Ground State of a molecule, we suppose that the potential operator, even if non-local,
can be effectively restricted on wavefunctions that can be considered as nonzero only on a
bounded domain. In other terms, a molecule can be associated to a ``scatterer'' that has a \emph{finite} 
spatial extension; if a state $\ket{\psi}$ is localized, it will be the same for the ket $\ket{\op V \psi}$, though 
perhaps in a larger region.
For a computational analysis of the problem, this property is fundamental in the choice of the numerical treatment as we know that
computational basis sets 
that are tailored to express asymptotically vanishing wavefunctions are better suited for the solution of the problem.

@@

It can be seen that for each choice of $p$, the value $\omega = |\eps_p|$ represents the threshold level 
that govern the localization properties of the corresponding fluctuation state.
As a \emph{stringent} hypothesis, let us now assume that the ket $\ket{s_p^\Phi(\omega)}$ is localized, namely its evaluation can be
restricted (if suitable) on a bounded real-space domain.
%The localization character of fluctuation states can be assessed by following the same argument discussed in section \ref{SEopenSystem} for the eigenstates of the SE. In this case 
%the source term depends also from the perturbation added to the system so this analysis applies only to the class 
%of perturbing field for which $\ket{\op\Phi\psi_p}$ is localized, when projected
%in the $\r$-representation. 
Even when such locality argument holds true, for $\omega > |\eps_p|$, 
the Helmholtz kernel behaves as a spherical wave and  gives raise to a $\ket{f^\Phi_p(\omega)}$ which is \emph{delocalized} in the entire real-space domain.
Such delocalization is an intrinsic consequence of the value of $\omega$, and does not depend on the particular choice of 
$\op \Phi$.
Only for $\omega < |\eps_p|$ the kernel contains an exponential damping factor and $\ket{f^\Phi_p(\omega)}$ exhibits a bound-state behavior in the long range, 
once again provided that the above locality arguments on $\ket{s_p^\Phi(\omega)}$ -- hence on $\op V'$ and $\Phi$ -- are valid.

Such locality property of the fluctuation state can be directly confirmed for static perturbations at $\omega=0$.
Let us consider a static perturbing operator $\op\Phi_s$ that can be restricted to a finite domain. 
Any local operator, like \emph{e.g.} a static electric field $\mathbf F$, with associated scalar potential $\Phi_s(\r) = -\mathbf F \cdot \r$, would satisfy such property.
We consider the perturbed hamiltonian $\hnot[\dm_{\Phi_s}] + \op\Phi_s$, where $\dm_{\Phi_s}$ represents the GS density of the perturbed system, and assume that the 
perturbation strength ($|\mathbf F|$ in the above example) is sufficiently small to be consistent with a LR treatment of the problem. 
In this case, the occupied eigenfunction $\ket{\psi_p^{\Phi_s}}$ of the perturbed hamiltonian can be extracted with traditional GS techniques, 
as from the discussion around Eq.~\eqref{SEHelmholtz1}, they exhibit bound-state behavior.\footnote{The fulfilling of Eq.~\eqref{SEHelmholtz1} for energies below threshold
is not \textit{per se} a guarantee of a localized $\ket{\psi}$. With such equation we demonstrate that, \emph{in the case when} a localized
solution exists, this cannot be found for energies above threshold.}
\begin{figure}[t]
\includegraphics[scale=0.68]{CO_statPolvsBox.eps}
\caption{\label{co_alphaStatic}(Color online) Convergence of the statical polarizability (blue continuous line) and of the ground state permanent dipole (red dotted line) 
of the $CO$ molecule with respect to the size of the computational domain. The inset plot evidences the relative percentage convergence using the value associated to the largest 
box as reference.}
\end{figure}

@@

This fact has interesting implication in the evaluation of LR quantities. Indeed, writing the response 
density in terms of fluctuation states implies that 
\be\lb{LinearResponseFunctDef2}
\braket{\delta\op O}_\Phi(\omega) = 2\mathrm{Re} \left(\sum_p \bra{\psi_p}\op O\ket{f_p^\Phi(\omega)}\right) \;.
\ee
%so the computation of the linear response functional passes through the evaluation of the scalar products indicated in \eqref{LinearResponseFunctDef2}.
Let us suppose that also the observable of interest $\op O$ exhibit the locality properties of the potentials $ \op V$, $\op \Phi$, $\op V'$.
Once again, it would be enough to consider a local operator for this condition to be met.
We observe that there exists two 
distinct regimes, on the basis of value of $\omega$. 
The first one, which we will refer to as the \emph{below threshold} regime, is realized when
$\omega$ is lower than the ionization potential $|\eps_h|$.
In this case all the fluctuation states are below their threshold level and equation \eqref{LinearResponseFunctDef2} 
is expressed in terms of genuinely localized 
quantities. The quantities $\braket{\psi_p \op O| f_p^\Phi(\omega)}$, 
being scalar products of bound-state-like objects, may thus be evaluated with computational setups that
are similar to the ones usually employed for GS calculations.
This consideration proves why, when a given computer code employs established GS numerical techniques, 
\emph{it is much easier to converge LR quantities for $\omega$ below IP}.
The convergence in this regime is ``strong'', in the sense that the basis functions employed enable to express 
with arbitrary precision \emph{both} the states $\ket{\op O \psi_p}$ and $\ket{f_p^\Phi(\omega)}$.
This is of course true even when the above states are only implicitly expressed by the employed LR treatment,
 as e.g. in the case of Krylov spaces generated from the states $\ket{\op\Phi(\omega) \psi_p}$ \cite{baroni2006,baroni2008,linlinKPM}.

@@

\subsection{The analytic structure of linear susceptibility}
Being themselves transverse operators, without loss of generality we can parametrize excitations by employing  
the \emph{excited states} $\ket{\phi^a_p}$ and $\bra{\chi^a_p}$ that, by construction, are orthogonal to the occupied orbitals
$\ket{\psi_p}$ and satisfy the normalization condition % and the orthonormalization condition \eqref{orthoExcitatioOpDef1} implies 
\be\lb{ExcitedStateOrthNormDef1}
\sum_p \left(\brket{\phi_p^a}{\phi_p^b} - \brket{\chi_p^b}{\chi_p^a}\right) = \delta_{ab} \;. 
\ee
In this way we can parametrize the excitation operators of the system as:
\begin{align}\lb{ExcitationOperatorsDef2}
\op E_a &= \sum_p \left( \ket{\phi^a_p}\bra{\psi_p} + \ket{\psi_p} \bra{\chi^a_p}\right)\;, \nn \\
\op{\tilde E}_a = \commutator{\dmnot}{\op E^t_a} &= \sum_p \left( \ket{\psi_p}\bra{\phi^a_p} - \ket{\chi^a_p} \bra{\psi_p}\right)\;.
\end{align}
Each excitation ``mode'' of the system $\op E_a$ with energy $\Omega_a$, can then be associated to a set of states $\{\phi^a_p,\chi^a_p\}_{p \in a}$,
entirely defined in the unoccupied subspace. For each choice of $p,a$, these objects represent, respectively, the state in which the occupied state $\ket{\psi_p}$
is excited -- or from whom it decays -- when the system is subject to the perturbation $\Phi_a \equiv \commutator{\dmnot}{\op E_a}$.
We denote with the shorthand $p \in a$ the fact that the sum of Eq.~\eqref{ExcitationOperatorsDef2}
is in general restricted only to a $a$-dependent subset of the occupied orbitals, depending on the particular symmetry 
of the excitation.

% We observe that right eigenvalues of $\Liouv$ can be written in terms of the left ones due to the properties 
% of the Liouvillian described in appendix \ref{LiouvillianAction}. 
% By applying equations \eqref{LiouZeroDef1},\eqref{CouplDef1} and \eqref{LiouvillianRightActionDef1} presented in 
% appendix \ref{LiouvillianAction} to the dyadic representations of the excitation 
% operators, 
By applying the equations \eqref{ExcitationOperatorsDef1} to the above parametrization
of the excitations, we obtain the following Sternheimer-like equations for the excited states:
\begin{align}\lb{ExcitationOperatorsDef3}
&\left[\Omega_a-(\hnot - \eps_p)\right] \ket{\phi_p^a} = \op Q_0 \op V'[\op E_a] \ket{\psi_p} \;, \nn\\
&\left[-\Omega_a-(\hnot - \eps_p)\right]\ket{\chi_p^a} = \op Q_0\op V'[\op E^t_a]\ket{\psi_p}  \;.
\end{align}
By writing these equations in a basis set of virtual states we obtain the well-known Casida's eigenvalue equation
associated to the excitation energy $\Omega_a$ (see appendix \ref{casida}).
The excitation spectrum is symmetric w.r.t the inversion of the eigenvalues 
$\Omega_a \rightarrow -\Omega_a$ and the associated excitation with negative energy are described by the transpose pair $\{\chi^a_p,\phi^a_p\}$.
We concentrate our analysis on the sector of positive energies, i.e. $\Omega_a > 0$.

With the same spirit of the previous sections, we present a formal solution of \eqref{ExcitationOperatorsDef3} as follows 
\begin{align}
\ket{\phi^a_p} &= \op G_H(\Omega_a+\eps_p)\left(\op V \ket{\phi^a_p} + \ket{s_{\phi}} \right) \;, \nn \\
\ket{\chi^a_p} &= \op G_H(-\Omega_a+\eps_p)\left( \op V \ket{\chi^a_p} + \ket{s_{\chi}} \right)\;,
\end{align}
where the source terms $s_\phi$ and $s_{\chi}$ are defined as:
\begin{align}
 \ket{s_{\phi}} &=  \op Q_0 \op V'[\op E_a] \ket{\psi_p} \\
 \ket{s_{\chi}} &=  \op Q_0 \op V'[\op E^t_a] \ket{\psi_p}\;.
\end{align}
We first point out that the Helmholtz kernel associated to the $\ket{\chi^a_p}$ state 
contains an exponential damping factor for \emph{all} the positive values of $\Omega_a$.
The asymptotic long-range behaviour of $\ket{\chi^a_p}$ thus depends only on the locality properties of 
$\ket{s_{\chi}}$, which in turn is related to the behaviour of the operator $\op V'[\op E^t_a]$.
% As in the previous section, we assume the hypothesis that such operator is such that
% the state $\op Q_0 \op V'[\op E^t_a] \ket{\psi_p}$ is localized, at least when \emph{both}
% the states $\phi_p^a$ $\chi_p^a$ have bound-state-like behaviour.

On the other hand, the long-range behaviour of $\op G_H(\Omega_a+\eps_p)$ depends from 
$\Omega_a$, with a threshold of value $\Omega_a=|\eps_p|$. 
For energies above this value, the behaviour of the state $\ket{\phi^a_p}$ is 
delocalized regardless of the particular nature of the operators $\op V$ and $\op V'$.
This means that, if an excitation operator $\op E_a$ has an energy such that
$\Omega_a>|\eps_p|$  for \emph{at least one} of the $p \in a$ defining the excitation,
such operator cannot be parametrized by employing localized states only.

@@

Instead, the second sector contains the dense set of delocalized excitations with an associated continuum spectrum. These excitations behave as the unbound virtual orbitals analyzed in \cite{boffi2016} 
and are deeply affected by the computational setup used to represent them. In particular, the presence of a finite computational domain implies a fictitious discretization of the excitation levels and 
the associated spectrum exhibit an explicit dependence from the size of the domain. 
Excitations belonging to this sector thus lose their intrinsic physical meaning and have to be understood simply as
a representation of the density of states of the Liouvillian superoperator.

@@

We have seen in Sec.~\ref{FluctuationState} that 
for values of $\omega$ below the first ionization potential we expect
$\ket{f_p^\Phi(\omega)}$ to have bound-state like behavior.
For this to be true, the contribution from the delocalized excited states defining
$\ket{f_p^\Phi(\omega)}_C$ should resum into a localized, asymptotically vanishing state.
Indeed, there is no particular reason to believe that $\ket{f_p^\Phi(\omega)}_C$ should be
zero. The linear independence of the $\ket{w_p^a}$ states would imply that 
$\varphi_p^a(\op \Phi,\omega)=0$ for any $a$ labelling the continuum excitations,
which is too strong a condition for a general choice of $\op \Phi$.
\footnote{The plot in Fig.~\ref{co_AlphaExc}
show this evidence even in the simple case when $\op \Phi$ is the dipole operator.}.

This means that, contrary to GS, \emph{for LR calculation,
even when the response density is localized, the contribution coming from the 
continuum sector of the excitations can never be neglected}. 
We are not referring here to the continuum eigenstates of the 
unperturbed GS hamiltonian, %which is not in itself important for expressing the linear susceptivity, 
but to the continuum of excitations expressed as eigenstate of the 
Liouvillian superoperator of Eq.~\eqref{Liouvillian}, needed to 
express the spectral representation of the linear susceptibility.
We \emph{cannot} exclude that a fluctuation state, even if genuinely localized, will have a nonzero projection
in the essential part of the Liouvillian spectrum.
